---
title: "Introduction to RMarkdown Tutorial"
author: Add your name here
output: 
  html_document:
    number_sections: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document will describe the basic setup for using RMarkdown, and how
it can be used to mix writing and statistical analysis.

# Confirm you can knit this

First, before making any changes, simply press `Knit` above, to create
the "knitted" html output of this document. It will not contain any
code, just the text directions here.

# Perform a simple calculation

In the empty code chunk below, compute the average of the numbers
1,2,3,4,5. Your output, when you knit the document, should show the code
you use to do this **and the answer**.

```{r}

mean(c(1,2,3,4,5))

```

# Write a comment in your code chunk

Now, instead compute the median of the same 5 numbers. In the code
chunk, write a brief comment, starting with `#`, explaining what the
median is. Notice how this is printed when you knit the file.

```{r}
# median is...

```

# Create a plot

In the below code, use the `rnorm` function to generate 100 random draws
from a normal distribution with mean 0 and variance 1. Save these in an
object called `x`. Then run `hist(x)`. Use comments (`#`) to describe
the plot that is created, and what happens if you re-knit this to the
plot.

```{r}


x <- rnorm(n=100, mean=0, sd=1)
hist(x)
# this plot describes...

```

# Add your name and save as a pdf

Finally, modify the second line in this document, after `author` to
include your own name. Then, after knitting this, open the created html
document in a browser and save it as a pdf. This is how you will need to
create your homework and project submissions.
